---
title: "Intro to RMarkdown"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Headers

#### Deeper level header

Write some text. Rmarkdown is really cool.

This is a new line. We need to include a black line to creat a new line.

-   apple
-   banana
-   kiwi

# Level 1 header

## Level 2 header

*writing in italic; does it work?*

1.  **apple**

2.  banana

3.  kiwi

    -   Item 3a
    -   Item 3b

[Artic Data Center](http://articdata.io)

[@oke2020]

| Fruit | Color | Deliciousness |
|-------|-------|---------------|
| apple | red   | no            |
| apple | green | yes           |
| pear  | green | yes           |

: Fruit Tastiness


To insert a new code chunk we have **two** options:

1. Insert code chunk icon/dropdown
2. Ctrl + Atl + i



```{r}
x <- 3 * 3 
x
```


It is better to include all the packages at the top of our Markdown

```{r}
library(readr)
```


## Introduction

* read in and reference data source

*echo shows the code in the Markdown, message shows the output message*

```{r read-data, message = FALSE}
bg_chem <- read_csv("BGchem2008data.csv")
```


## Analysis

* calculate basic summary stats
  - mean NO3
  - mean NO2
  - mean NH4
  - mean P

```{r basic summary stats}
NO3 <- mean(bg_chem$NO3)
NO2 <- mean(bg_chem$NO2)
NH4 <- mean(bg_chem$NH4)
P <- mean(bg_chem$P)
```

* calculate mean redfield ratio

```{r}
red <- (NO3 + NO2 + NH4) / P
```

We calculated the Redfield ratio to be `r round(red, 2)`


* plot redfield ratio

```{r, echo = FALSE}
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```

How do I decide when to make a new chunk?

*We should use one chunk for one "thing"*

*One plot per chunk is ideal*

*If we are using too much comments it's a sign*

**How to convert R chunks into a R script?**

https://bookdown.org/yihui/rmarkdown-cookbook/purl.html

