---
title: "Session 12 exercise"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

```{r, message = FALSE}
library(rvest)
library(readr)
library(dplyr)
library(janitor)
library(ggplot2)
```

```{r}
webpage <- read_html("https://www.pwrc.usgs.gov/bbl/manual/speclist.cfm")

tbls <- html_nodes(webpage, "table") %>% 
    html_table(fill = TRUE)

species <- tbls[[1]] %>% 
    clean_names() %>% 
    select(alpha_code, common_name) %>% 
    mutate(alpha_code = tolower(alpha_code))
```

## My theme

```{r}
my_theme <- theme_classic() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

## Loading data

```{r, message = FALSE}
nest_data <- read_csv("https://cn.dataone.org/cn/v2/resolve/urn:uuid:982bd2fc-4edf-4da7-96ef-0d11b853102d")

predator_survey <- read_csv("https://cn.dataone.org/cn/v2/resolve/urn:uuid:9ffec04c-7e2d-41dd-9e88-b6c2e8c4375e")

```

## Creating a function that will translate species codes into common names

```{r}
#' Function to add common name to data.frame according to the BBL list of species codes
#'
#' @param df A data frame containing BBL species codes in column `species`
#' @param species A data frame defining BBL species codes with columns `alpha_code` and `common_name`
#'
#' @return  data frame with original data df, plus the common name of species
#'
#' @examples
translate_species <- function(df, species) {
  
  return_df <- left_join(df, species, by = c("species" = "alpha_code"))
  
  
  return(return_df)
}

```

```{r}
new_nest_data <- translate_species(nest_data, species)
new_predator_survey <- translate_species(predator_survey, species)
```

## Calculate total number of predators by year and species, and plot the result

```{r}
n_predator <- new_predator_survey %>% 
  group_by(year, common_name) %>%
  summarise(pred_count = sum(count, na.rm = T))

ggplot(n_predator, aes(x = year, y = pred_count, color = common_name)) +
  geom_point()+
  geom_line()+
  ylab("Number of individuals") +
  xlab("Year") +
  ggtitle("Predator species") +
  my_theme

ggplot(n_predator, aes(x = common_name, y = pred_count, fill = common_name)) +
  geom_col()+
  ylab("Number of individuals") +
  xlab("Species names") +
  ggtitle("Predator species") +
  my_theme
```


## Calculate total number of eggs predated by year and species

```{r}
n_eggs_predated <- new_nest_data %>% 
  group_by(year, common_name) %>%
  summarise(n_eggs = sum(number_eggs_predated, na.rm = T))

ggplot(n_eggs_predated, aes(x = year, y = n_eggs, color = common_name)) +
  geom_point()+
  geom_line()+
  ylab("Number of eggs predated") +
  xlab("Year") +
  ggtitle("Eggs predated") +
  my_theme

ggplot(n_eggs_predated, aes(x = common_name, y = n_eggs, fill = common_name)) +
  geom_col()+
  ylab("Number of eggs predated") +
  xlab("Species names") +
  ggtitle("Eggs predated") +
  my_theme
```

## Calculate total number of predators by year, join to summarized egg predation table, and plot the result

```{r}
predator_by_year <- new_predator_survey %>% 
  group_by(year) %>%
  summarise(pred_count = sum(count, na.rm = T))

n_eggs_predated <- left_join(n_eggs_predated, predator_by_year, by = "year")

ggplot(n_eggs_predated, aes(x = n, y = n_eggs)) +
  geom_point()+
  ylab("Number of eggs predated") +
  xlab("Number of predators") +
  my_theme

ggplot(n_eggs_predated, aes(x = n, y = n_eggs)) +
  geom_point()+
  facet_wrap(~ common_name, scales = "free_y", ncol = 3) +
  ylab("Number of eggs predated") +
  xlab("Number of predators") +
  my_theme

```

## What species of predator is the most abundant and has this changed through time?

```{r}

```

